---
title: "Functional_Data_Example"
author: "Angela Yu"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Functional_Data_Example}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  message   = FALSE,
  warning   = FALSE,
  fig.width = 6
)
```

The example here is to use pupil functional data to construct the simultaneous outer and inner confidence region from simultaneous confidence bands (SCB) using Function-on-Scalar Regression (FoSR).

The pupil dataset contains repeated measures of percent change over time for multiple subjects under two user categories (use: 1 and no use: 0). It contains both user and non-user groups, multiple time points, and several covariates, including age, gender, BMI, and alcohol consumption.

```{r load pupil data, echo = TRUE}
library(SCoRES)
library(mgcv)
data(pupil)
```

Before calculating the SCBs, we first process pupil data by fitting a mean GAM model, extracting residuals and performing FPCA using `SCoRES::prepare_pupil_fpca()`, the function will return an enhanced dataset includes the FPCA-derived basis scores (Phi1, Phi2, Phi3, Phi4) for Function-on-Scalar Regression (FoSR) analysis.

Following the FPCA-based data augmentation, we fit a FoSR model using `mgcv::bam()`, which allows efficient estimation of Generalized Additive Mixed Models (GAMMs). The model formula is designed to capture both population-level smooth trends and subject-specific functional variation.

The response here is percent_change. Time variable t is seconds. And the covariate is use, which is a binary variable (1:use, 0:no use).

The function-on-scalar regression model is

$$
Y_i(t) = \beta_0(t) + \beta_1(t) X_{i1} + b_i(t) + \epsilon_i(t),
$$

```{r fit gam-fpca model, echo = TRUE}
pupil_fpca <- SCoRES::prepare_pupil_fpca(pupil)
fosr_mod <- mgcv::bam(percent_change ~ s(seconds, k=30, bs="cr") +
            s(seconds, by = use, k=30, bs = "cr") +
            s(id, by = Phi1, bs="re") +
            s(id, by = Phi2, bs="re") +
            s(id, by = Phi3, bs="re") +
            s(id, by = Phi4, bs="re"),
            method = "fREML", data = pupil_fpca, discrete = TRUE)
```

After obtaining the FoSR object `fosr_mod`, simultaneous confidence bands (SCB) can be constructed though `SCoRES::SCB_functional_outcome` using pre-specified methods.
The SCoRES package provides two options for calculating simultaneous confidence bands (SCBs), specified via the method argument:
`cma`: Correlation and Multiplicity Adjusted (CMA) confidence bands
via parametric approach.
`multiplier`: Dense confidence bands via Multiplier-t Bootstrap method.
Use `subset` to specify the names of grouping variables to analyze. The input data should have numerical binary 0/1 values for all scalar group variables. Here, we analyze the user group by specifying `subset = c("use = 1")`.
Use `fitted` to specify the object for SCB estimation. If `fitted = TRUE`, `SCoRES::SCB_functional_outcome` will construct the SCB for the fitted mean outcome function. If `fitted = FALSE`, `SCoRES::SCB_functional_outcome` will construct the SCB for the fitted parameter function.
Use est_mean to specify the point estimate used for constructing SCBs in SCoRES::SCB_functional_outcome. This argument is for `multiplier`
If est_mean = `TRUE`, the multiplier bootstrap will be performed based on the fitted mean outcome function.
If est_mean = `FALSE`, the multiplier bootstrap will be performed based on the sample mean.
For `multiplier` option, users need to provide a functional regression object through argument object. If object = `NULL`, the function will only output the SCB for an overall mean outcome function regardless of the group specified.

Here, we estimated SCBs using both options separately for the mean outcome Y(t) of user's group:
$$
\hat{f}(t) = E[Y(t) | X_{1} = 1]= \beta_0(t) + \beta_1(t),
$$
where:

- \( Y_i(t) \) is a functional outcome (percent_change)
- \( X_{i1} \) is a scalar covariate (use)

```{r pupil SCB, echo = TRUE}
# CMA approach
results_pupil_cma <- SCoRES::SCB_functional_outcome(
                                          data_df = pupil,
                                          object = fosr_mod, 
                                          method = "cma",
                                          fitted = TRUE,
                                          est_mean = TRUE, 
                                          alpha = 0.05, 
                                          outcome = "percent_change",
                                          domain = "seconds", 
                                          subset = c("use = 1"), 
                                          id = "id")

#CMA approach for parameter function
results_pupil_cma_para <- SCoRES::SCB_functional_outcome(
                                          data_df = pupil,
                                          object = fosr_mod, 
                                          method = "cma",
                                          fitted = FALSE,
                                          est_mean = TRUE, 
                                          alpha = 0.05, 
                                          outcome = "percent_change",
                                          domain = "seconds", 
                                          subset = c("use = 1"), 
                                          id = "id")

# Multiplier-t Bootstrap
results_pupil_multiplier <- SCoRES::SCB_functional_outcome(
                                          data_df = pupil,
                                          object = fosr_mod, 
                                          method = "multiplier",
                                          fitted = TRUE,
                                          est_mean = TRUE, 
                                          alpha = 0.05, 
                                          outcome = "percent_change",
                                          domain = "seconds", 
                                          subset = c("use = 1"), 
                                          id = "id")

results_pupil_multiplier_para <- SCoRES::SCB_functional_outcome(
                                          data_df = pupil,
                                          object = fosr_mod, 
                                          method = "multiplier",
                                          fitted = FALSE,
                                          est_mean = TRUE, 
                                          alpha = 0.05, 
                                          outcome = "percent_change",
                                          domain = "seconds", 
                                          subset = c("use = 1"), 
                                          id = "id")

# Multiplier-t Bootstrap (use sample mean)
results_pupil_multiplier_sample <- SCoRES::SCB_functional_outcome(
                                          data_df = pupil,
                                          object = fosr_mod, 
                                          method = "multiplier",
                                          fitted = TRUE,
                                          est_mean = FALSE, 
                                          alpha = 0.05, 
                                          outcome = "percent_change",
                                          domain = "seconds", 
                                          subset = c("use = 1"), 
                                          id = "id")

results_pupil_multiplier_overall <- SCoRES::SCB_functional_outcome(
                                          data_df = pupil,
                                          method = "multiplier",
                                          fitted = TRUE,
                                          est_mean = FALSE, 
                                          alpha = 0.05, 
                                          outcome = "percent_change",
                                          domain = "seconds", 
                                          subset = c("use = 1"), 
                                          id = "id")
```

In the multiplier bootstrap procedure, `SCoRES` supports three types of multiplier distributions, which is specified by `weights`:

- `"rademacher"`: $g_i \in \{-1, +1\}$ with equal probability
- `"gaussian"`: $g_i \sim \mathcal{N}(0, 1)$
- `"mammen"`: A two-point distribution with mean zero and variance one (see Mammen, 1993)

Default is `rademacher`.

Two options are available for estimating the standard error 
$\hat{\epsilon}_N^*(s_j)$, which is specified by `method_SD`:

- "regular" (empirical standard error based on residuals):

- "t" (bootstrap second moment-based estimator):

Default is `t`.

For mathematical details, see vignette Methods.
  
The code below visualizes the simultaneous outer and inner region derived from SCB results using the `SCoRES::plot_cs()` function. The `results` object is first converted to a tibble for easier manipulation.

The `levels = c(-18, -20, -22, -24)` argument specifies a set of thresholds, and `SCoRES::plot_cs()` function estimates multiple inverse upper excursion sets corresponding to these thresholds, and plot the estimated inverse region, the inner confidence region, and the outer confidence region. 

```{r pupil_plot_cs_cma, echo = TRUE}
results_pupil_cma <- tibble::as_tibble(results_pupil_cma)
plot_cs(results_pupil_cma,
        levels = c(-18, -20, -22, -24), 
        x = results_pupil_cma$domain, 
        mu_hat = results_pupil_cma$mu_hat, 
        xlab = "Seconds", 
        ylab = "Percent_Outcome", 
        level_label = T, 
        min.size = 40, 
        palette = "Spectral", 
        color_level_label = "black")
```

The plot demonstrate how to use SCB to find regions of s where the estimated mean is greater than or equal to the four levels -18, -20, -22 and -24 for ccds data. The gray shaded area is the 95% SCB, the solid black line is the estimated mean. The red horizontal line shows the inner confidence region (where the lower SCB is greater than the corresponding level) that are contained in the estimated inverse region represented by the union of the yellow and red horizontal line (where the estimated mean is greater than the corresponding levels); the outer confidence region are the union of the blue, yellow and red line (where the upper SCB is greater than the corresponding levels) and contain both the estimated inverse region and the inner confidence region.

The next plot shows the SCB for the coefficient function for the use group.

```{r ccds_plot_cs_pupil_para, echo = TRUE}
results_pupil_cma_para <- tibble::as_tibble(results_pupil_cma_para)
plot_cs(results_pupil_cma_para,
        levels = c(4.5, 5, 5.5, 6), 
        x = results_pupil_cma_para$domain, 
        mu_hat = results_pupil_cma_para$mu_hat, 
        xlab = "Seconds", 
        ylab = "Percent_Outcome", 
        level_label = T, 
        min.size = 40, 
        palette = "Spectral", 
        color_level_label = "black")
```

The following plots show the results of multiplier bootstrap. The only difference between these two plots is that, the solid black line in the first one represents the estimated mean fitted through functional regresson object, while the second represents the sample mean. All possible NA's are imputed using `refund::fpca.face()`.

```{r pupil_plot_cs_multiplier, echo = TRUE}
results_pupil_multiplier <- tibble::as_tibble(results_pupil_multiplier)
plot_cs(results_pupil_multiplier,
        levels = c(-18, -20, -22, -24), 
        x = results_pupil_multiplier$domain, 
        mu_hat = results_pupil_multiplier$mu_hat, 
        xlab = "Seconds", 
        ylab = "Percent_Outcome",
        level_label = T, 
        min.size = 40, 
        palette = "Spectral", 
        color_level_label = "black")
```

```{r pupil_plot_cs_multiplier_para, echo = TRUE}
results_pupil_multiplier_para <- tibble::as_tibble(results_pupil_multiplier_para)
plot_cs(results_pupil_multiplier_para,
        levels = c(4.5, 5, 5.5, 6), 
        x = results_pupil_multiplier_para$domain, 
        mu_hat = results_pupil_multiplier_para$mu_hat, 
        xlab = "Seconds", 
        ylab = "Percent_Outcome",
        level_label = T, 
        min.size = 40, 
        palette = "Spectral", 
        color_level_label = "black")
```

```{r pupil_plot_cs_multiplier_sample, echo = TRUE}
results_pupil_multiplier_sample <- tibble::as_tibble(results_pupil_multiplier_sample)
plot_cs(results_pupil_multiplier_sample,
        levels = c(-18, -20, -22, -24), 
        x = results_pupil_multiplier_sample$domain, 
        mu_hat = results_pupil_multiplier_sample$mu_hat, 
        xlab = "Seconds", 
        ylab = "Percent_Outcome", 
        level_label = T, 
        min.size = 40, 
        palette = "Spectral", 
        color_level_label = "black")
```

```{r pupil_plot_cs_multiplier_overall, echo = TRUE}
results_pupil_multiplier_overall <- tibble::as_tibble(results_pupil_multiplier_overall)
plot_cs(results_pupil_multiplier_overall,
        levels = c(-18, -20, -22, -24), 
        x = results_pupil_multiplier_overall$domain, 
        mu_hat = results_pupil_multiplier_overall$mu_hat, 
        xlab = "Seconds", 
        ylab = "Percent_Outcome", 
        level_label = T, 
        min.size = 40, 
        palette = "Spectral", 
        color_level_label = "black")
```

To further illustrate the power of `SCoRES::SCB_functional_outcome` for constructing SCBs for multiple group variables, we analyze the 40-years-old male user group by specifying
`group_name = c("use", "age", "gender")` and `group_value = c(1, 40, 0)`. We set `fitted = TRUE` and `est_mean= TRUE`.

The SCBs are estimated using both options separately for the mean outcome Y(t) of the specified group:
$$
\hat{f}(t) = E[Y(t) | X_{1} = 1, X_{2} = 40, X_{3} = 0]= \beta_0(t) + \beta_1(t) + 40*\beta_2(t),
$$
where:

- \( Y_i(t) \) is a functional outcome (percent_change)
- \( X_{1} \) is a scalar covariate (use)
- \( X_{2} \) is a continuous covariate (age)
- \( X_{3} \) is a scalar covariate (gender)

```{r pupil SCB 2, echo = TRUE}
# CMA approach
results_pupil_cma <- SCoRES::SCB_functional_outcome(
                                          data_df = pupil,
                                          object = fosr_mod, 
                                          method = "cma", 
                                          fitted = TRUE,
                                          est_mean = TRUE, 
                                          alpha = 0.05, 
                                          outcome = "percent_change", 
                                          domain = "seconds", 
                                          subset = c("use = 1", 
                                                     "age = 40", 
                                                     "gender = 0"),  
                                          id = "id")

# Multiplier-t Bootstrap
results_pupil_multiplier <- SCoRES::SCB_functional_outcome(
                                          data_df = pupil,
                                          object = fosr_mod,
                                          fitted = TRUE,
                                          method = "multiplier", 
                                          est_mean = TRUE, 
                                          alpha = 0.05, 
                                          outcome = "percent_change", 
                                          domain = "seconds", 
                                          subset = c("use = 1", 
                                                     "age = 40", 
                                                     "gender = 0"), 
                                          id = "id")
```

The following plots show the SCBs and inverse SCBs for the mean outcome function. The first one is from CMA approach and the second is from multiplier bootstrap.

```{r pupil_plot_cs_cma 2, echo = TRUE}
results_pupil_cma <- tibble::as_tibble(results_pupil_cma)
plot_cs(results_pupil_cma,
        levels = c(-18, -20, -22, -24), 
        x = results_pupil_cma$domain, 
        mu_hat = results_pupil_cma$mu_hat, 
        xlab = "Seconds", 
        ylab = "Percent_Outcome",
        level_label = T, 
        min.size = 40, 
        palette = "Spectral", 
        color_level_label = "black")
```

```{r pupil_plot_cs_multiplier 2, echo = TRUE}
results_pupil_multiplier <- tibble::as_tibble(results_pupil_multiplier)
plot_cs(results_pupil_multiplier,
        levels = c(-18, -20, -22, -24), 
        x = results_pupil_multiplier$domain, 
        mu_hat = results_pupil_multiplier$mu_hat, 
        xlab = "Seconds", 
        ylab = "Percent_Outcome",
        level_label = T, 
        min.size = 40, 
        palette = "Spectral", 
        color_level_label = "black")
```
