---
title: "Geographic_Data_Example"
author: "Angela Yu"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Geographic_Data_Example}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  message   = FALSE,
  warning   = FALSE,
  fig.width = 10,
  fig_height = 10
)
```

The following example is for displaying the confidence region for the increase of the mean summer temperature (June–August) in North America between the 20th and 21st centuries according to the specific climate model analyzed in Sommerfeld et al.(2018). 

`SCoRES::SCB_gls_geospatial()` function uses multiplier bootstrap algorithm to construct the SCB for a specified linear combination of coefficients across a geographical area. The argument `sp_list` specifies a list containing the spatial coordinates and the observations on each spot. `data_fit` specifies a data frame containing the design matrix used to fit the spatial generalized least square model (notice intercept should also be included inside the data frame), while `w` defines the weights for the linear combination of coefficients on which the simultaneous confidence bands (SCB) are constructed. Use the `correlation` argument to specify the correlation structure for model fitting. If the observations are grouped, a grouping factor can be specified using the `group` argument, and a covariance array `V` may also be provided to supply a known covariance matrix for each spot. `mask` is an optional logical matrix indicating spatial locations to include in the SCB, with excluded areas (e.g., water) set to 0 or NA.

```{r climate_data-est_SCB}
library(patchwork)
library(nlme)
library(SCoRES)
library(ggplot2)

# Load data
data(climate_data)
start_time <- Sys.time()

# construct confidence region for the increase of the mean temperature (June-August) in North America between the 20th and 21st centuries
temp = SCB_gls_geospatial(sp_list = climate_data$Z, 
                       level = 2, 
                       data_fit = climate_data$X,
                       w = c(1,0,0,0),
                       correlation = climate_data$correlation,
                       mask = climate_data$mask, 
                       alpha = 0.1)

end_time <- Sys.time()
time_taken <- end_time - start_time
print(time_taken)
```

Heat maps show the estimate of the mean temperature difference. The first row displays the contours of the outer confidence region, estimated inverse region, and the inner confidence region, for various levels. The three plots in the second row display the confidence region for the inverse sets, where the estimated mean difference is greater or equal to the individual level 1.5, 2.0, or 2.5 respectively. In the second
 row, the blue line is the contour of the outer confidence region, the green line is the contour of the estimated inverse region and the red line is the contour of the inner confidence region.
 
```{r climate_data-plot_cs}
par(mfrow = c(2, 3), mar = c(3, 3, 2, 1)) 
p2 = plot_cs(list(scb_up = temp$scb_up, scb_low = temp$scb_low), levels = c(1.5, 2,2.5,3), x = temp$x, y = temp$y, mu_hat = temp$mu_hat, xlab = "Longitude", ylab = "Latitude", level_label = T, min.size = 40, palette = "Spectral", color_level_label = "black")
p1 = plot_cs(list(scb_up = temp$scb_up, scb_low = temp$scb_low), levels = c(1.5,2,2.5), x = temp$x, y = temp$y, mu_hat = temp$mu_hat, xlab = "Longitude", ylab = "Latitude",together = F)
p = p2/p1
p
```


