---
title: "Linear_Model_Example"
author: "Angela Yu"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 2
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Linear_Model_Example}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
encoding: UTF-8
---

The example here is to use simulated data to construct the inverse confidence sets (CS) from simultaneous confidence bands (SCB) using linear regression. 

`invSCI::SCB_linear_outcome()` function use a non-parametric bootstrap algorithm to construct the SCB in linear regression. The argument `df_fit` specifies a data frame containing the training design matrix used to fit the linear model, while `grid_df` contains the test set design matrix for constructing SCB. Use argument `model` to specify the formula used for fitting the linear model.

```{r simulate data_estimate SCI_linear, echo = TRUE}
library(invSCI)
set.seed(262)
# generate simulated data
x1 <- rnorm(100)
x2 <- rnorm(100)
epsilon <- rnorm(100,0,sqrt(2))
y <- -1 + x1 + 0.5 * x1^2 - 1.1 * x1^3 - 0.5 * x2 + 0.8 * x2^2 - 1.1 * x2^3 + epsilon
df <- data.frame(x1 = x1, x2 = x2, y = y)
grid <- data.frame(x1 = seq(-1, 1, length.out = 100), x2 = seq(-1, 1, length.out = 100))
# fit the linear regression model and obtain the SCB for y
model <- "y ~ x1 + I(x1^2) + I(x1^3) + x2 + I(x2^2) + I(x2^3)"
results <- SCB_linear_outcome(df_fit = df, model = model, grid_df = grid)
```

The `levels = c(-0.3, 0, 0.3)` argument specifies a set of thresholds, and `invSCI::plot_cs()` function estimates multiple inverse upper excursion sets corresponding to these thresholds, and plot the estimated inverse set, the inner confidence set, and the outer confidence set.

```{r simulate_plot_cs_linear, echo = TRUE}
results <- tibble::as_tibble(results)
suppressWarnings(plot_cs(results,
        levels = c(-0.3, 0, 0.3), 
        x = seq(-1, 1, length.out = 100), 
        mu_hat = results$Mean, 
        xlab = "", 
        ylab = "", 
        level_label = T, 
        min.size = 40, 
        palette = "Spectral", 
        color_level_label = "black"))
```

In addition to linear regression, `invSCI` also  provides`invSCI::SCB_logistic_outcome()` for estimating the SCB for outcome of logistic regression.
```{r simulate data_estimate SCI_logistic, echo = TRUE}
# generate simulated data
x1 <- rnorm(100)
x2 <- rnorm(100)
epsilon <- rnorm(100,0,sqrt(2))
y <- -1 + x1 + 0.5 * x1^2 - 1.1 * x1^3 - 0.5 * x2 + 0.8 * x2^2 - 1.1 * x2^3 + epsilon
y <- expit(y)
df <- data.frame(x1 = x1, x2 = x2, y = y)
grid <- data.frame(x1 = seq(-1, 1, length.out = 100), x2 = seq(-1, 1, length.out = 100))
# fit the logistic regression model and obtain the SCB for y
model <- "y ~ x1 + I(x1^2) + I(x1^3) + x2 + I(x2^2) + I(x2^3)"
results <- SCB_logistic_outcome(df_fit = df, model = model, grid_df = grid)
```

Likewise, the `levels = c(0.3, 0.4, 0.5)` argument specifies a set of thresholds, and `invSCI::plot_cs()` function estimates multiple inverse upper excursion sets corresponding to these thresholds, and plot the estimated inverse set, the inner confidence set, and the outer confidence set.

```{r simulate_plot_cs_logistic, echo = TRUE}
results <- tibble::as_tibble(results)
plot_cs(results,
        levels = c(0.3, 0.4, 0.5), 
        x = seq(-1, 1, length.out = 100), 
        mu_hat = results$Mean, 
        xlab = "", 
        ylab = "", 
        level_label = T, 
        min.size = 40, 
        palette = "Spectral", 
        color_level_label = "black")
```

Besides, `invSCI::SCB_regression_coef` can estimate the SCB for every coefficient in the linear/logistic model.

```{r simulate data_estimate SCI_coef, echo = TRUE}
library(MASS)
# generate simulated data
M <- 50
rho <- 0.4
n <- 500
beta <- rnorm(M, mean = 0, sd = 1)
Sigma <- outer(1:M, 1:M, function(i, j) rho^abs(i - j))
X <- MASS::mvrnorm(n = n, mu = rep(0, M), Sigma = Sigma)
epsilon <- rnorm(n, mean = 0, sd = 1)
y <- X %*% beta + epsilon
df <- as.data.frame(X)
names(df) <- paste0("Beta", 1:M)
df$y <- as.vector(y)
# fit the linear regression model and obtain the SCB for all betas
model <- "y ~ ."
results <- SCB_regression_coef(df, model)
```

Likewise, the `levels = c(-2, -0.5, -0.1, 0.2)` argument specifies a set of thresholds, and `invSCI::plot_cs()` function estimates multiple inverse upper excursion sets corresponding to these thresholds, and plot the estimated inverse set, the inner confidence set, and the outer confidence set.

```{r simulate_plot_cs_coef, echo = TRUE}
results <- tibble::as_tibble(results)
plot_cs(results,
        levels = c(-2, -0.5, -0.1, 0.2), 
        x = names(df), 
        mu_hat = results$Mean, 
        xlab = "", 
        ylab = "", 
        level_label = T, 
        min.size = 40, 
        palette = "Spectral", 
        color_level_label = "black")
```

